Unsupervised acquisition of comprehensive multiword lexicons \\ using competition in an $n$-gram lattice

### 2016-09-28 eacl2017.tex
Tweaks to diagrams and text

@@

In this paper we are attempting to identify \termdef{formulaic sequences}, following the terminology of Wray \shortcite{Wray02,Wray08}, who defines a formulaic sequence (FS) as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' In other words, a formulaic sequence shows signs of being lexicalized. In computational linguistics, the most common term used to describe multiword lexical units is \emph{multiword expression} or MWE \cite{Baldwin10}, but here we wish to make a principled distinction between at least somewhat non-compositional, strongly lexicalized MWEs and FS, a superset which includes MWEs but also (apparently) compositional linguistic formulas. This distinction is not a new one, for instance it exists to some extent in the MWE annotation of Schneider \cite{Schneider14a}, who distinguishes between weak (collocational) and strong (non-compositional) MWEs, but it is our contention is that separate, precise terminology is useful for research targeted at either class: we need not strain the concept of MWE to include items which do not require special semantics, nor are we included to disregard or minimize the larger formulaticity of language simply because it is not the dominant focus of MWE research. Many MWE researchers would defensibly balk at including in their MWE lexicons and corpus annotations (English) formulaic sequences such as \ex{there is something going on}, \ex{it is more important than ever to \dotts}, \ex{\dotts do not know what it is like to \dotts}, \ex{it would be great if \gap could}, \ex{there is no shortage of\dotts}, \ex{the rise and fall of\dotts}, \ex{now is not the time to\dotts}, \etc as well as tens of thousands of other such phrases which, along with less compositional, clearly MWE formulas like \ex{be worth \gap weight in gold}, are identified by the FS-extraction system presented here.

\begin{figure}[!t]
\center{\includegraphics[width=0.35\textwidth]{FS_diagram_1.pdf}}
\caption{Multiword Terminology}

\end{figure}

Figure 1 shows our conception of the terminology for multiword relationships, including FS, MWE, as well as collocations and (multiword) named entities. We use the term \textit{collocation} to refer to statistical co-occurrence without necessarily any lexicalization of the sort implied by our definition of FS. MWEs, again, are a non-compositional subset of FS, and they overlap with named entities. In this regard, we disagree with the MWE annotation schema of Schneider \cite{Schneider14a} which simply includes any named entity as an MWE, which, in the context of product reviews from the Web Treebank, means that that named entities form nearly 50\% of the MWEs types attested in that corpus, with many being phrases such as \textit{Tintman Nationwide Tints Ltd} or \textit{Hickory Furniture Mart} which are easily identifiable as named entities based on their form, but would not be considered part of the mental lexicon of even a tiny fraction of English speakers (not even, we might argue, most of these people who work for these businesses, they likely refer to them using shorter, informal names). The upshot of this choice that a significant portion of ``MWE'' identification in the dataset is in fact named entity recognition. Some well-known named entities are obviously good MWEs (\eg \textit{George Washington}, \textit{the United States}), but not the names of the vast majority of people, businesses, or products which have real-word referents; they do not belong in a lexicon of formulaic sequences which is intended to reflect the internal lexicon of (at least) a significant generalizable subset of English speakers, putting them outside the scope of this work.

@@

\begin{figure*}[!tb]
\center{\includegraphics{FS_diagram_2.png}}
\caption{A portion of an $n$-gram lattice. Solid lines indicate subsumption, dotted lines overlaps}
\end{figure*}[!tb]

>>>>

\newcommand{\figref}[2][]{Figure#1~\ref{#2}\xspace}

@@

In this paper we are attempting to identify \termdef{formulaic sequences}, following the terminology of Wray \shortcite{Wray02,Wray08}, wherein a formulaic sequence (FS) is defined as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' In other words, a formulaic sequence shows signs of being lexicalized. In computational linguistics, the most common term used to describe multiword lexical units is \emph{multiword expression} or ``MWE'' \cite{Sag02,Baldwin10}, but here we wish to make a principled distinction between at least somewhat non-compositional, strongly lexicalized MWEs and FS, a superset which includes MWEs but also (apparently) compositional linguistic formulas. This distinction is not a new one; it exists, for example, in the original paper of \newcite{Sag02} in the distinction between lexicalised and institutionalized phrases, and also to some extent in the MWE annotation of \newcite{Schneider14a}, who distinguish between weak (collocational) and strong (non-compositional) MWEs. It is our contention, however, that separate, precise terminology is useful for research targeted at either class: we need not strain the concept of MWE to include items which do not require special semantics, nor are we inclined to disregard the larger formulaticity of language simply because it is not the dominant focus of MWE research. Many MWE researchers would defensibly balk at including in their MWE lexicons and corpus annotations (English) formulaic sequences such as \ex{there is something going on}, \ex{it is more important than ever to \dotts}, \ex{\dotts do not know what it is like to \dotts}, \ex{there is no shortage of\dotts}, \ex{the rise and fall of\dotts}, \ex{now is not the time to\dotts}, \etc as well as tens of thousands of other such phrases which, along with less compositional MWEs like \ex{be worth \gap weight in gold}, come under the umbrella of FSs.

\begin{figure}[!t]
\center{\includegraphics[width=0.35\textwidth]{FS_diagram_1.pdf}}
\caption{Multiword Terminology}
\label{fig:terminology}
\end{figure}

\figref{fig:terminology} shows our conception of the relationship between FS, MWE and (multiword) named entities. %We use the term \termdef{collocation} to refer to statistical co-occurrence without necessarily any lexicalization of the sort implied by our definition of FS. 
% TJB: I'd suggest staying away from talking about collocations altogether: it's an overloaded term; many MWEs are *not* collocations under this definition; and it's not needed for this paper
MWEs are a non-compositional subset of FS, and they overlap with named entities (NEs). In this regard, we disagree with the MWE annotation schema of \newcite{Schneider14a} which includes any NE as an MWE, which, in the context of product reviews from the Web Treebank, means that that named entities form nearly 50\% of the MWEs types attested in that corpus, with many being phrases such as \ex{Tintman Nationwide Tints Ltd} or \ex{Hickory Furniture Mart} which are easily identifiable as NEs based on their form, but would not be considered part of the mental lexicon of even a tiny fraction of English speakers.% (not even, we might argue, most of these people who work for these businesses, they likely refer to them using shorter, informal names). 
% TJB: too subtle/controversial a point for this paper
The upshot of this choice that a significant portion of ``MWE'' identification in the dataset is in fact named entity recognition. Some well-known named entities are obviously good MWEs (\eg \ex{George Washington}, \ex{the United States}), but not the names of the vast majority of people, businesses, or products which have real-word referents; they do not belong in a lexicon of formulaic sequences which is intended to reflect the internal lexicon of (at least) a significant generalizable subset of English speakers, putting them outside the scope of this work.

@@

\begin{figure*}[!tb]
\center{\includegraphics[width=0.8\textwidth]{FS_diagram_2.pdf}}
\caption{A portion of an $n$-gram lattice. Solid lines indicate subsumption, dotted lines overlaps}
\end{figure*}
